---
# title: "vignette_CFSTRenD"
output: rmarkdown::html_vignette
 
vignette: >
  %\VignetteIndexEntry{vignette_CFSTRenD}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

### CFSTRenD R package aimed at streamlining the processing and analysis of tree-ring data. Key features include: 
###     1) manipulation and summarization of tree-ring data; 
###     2) data quality identification and classification; 
###     3) a flexible detrending and climate-growth model based on generalized additive mixed models (GAMM); 
###     4) model selection tool based on Akaike Information Criterion (AIC). 

### This vignette provides a precise guide to navigating key aspects of data processing, quality assessment, and modeling using the CFSTRenD package. Two datasets are used for demonstration:

###   1. samples69.csv: A subset of the project Douglas-fir Retrospective Monitoring, freely available from the online CFS-TRenD data collection ([https://ts.nrcan.gc.ca/treerings]). This dataset contains species Pseudotsuga menziesii (Douglas-fir) data from 20 sites, totaling 257 samples with ring-width measurements in a wide format.
###   2. clim69.csv: Annual winter frostdays (FFD) data corresponding to the same 20 sites included in samples69.csv.

### We start by loading the datasets:

```{r setup}
library(data.table)
library(CFSTRenD)
# read ring measurement
 samples69.o <- fread(system.file("extdata", "samples69.csv", package = "CFSTRenD"))
 # read climate
clim69 <- fread(system.file("extdata", "clim69.csv", package = "CFSTRenD"))

```

# 1. processing data compatible with CFS-TRenD format 
## this is to generates formatted tables aligning with CFS-TRenD structure, ensuring compatibility for data submission and integration.
```{r}

samples69.trt <- CFS_format(data = list(samples69.o, 39:140), out.csv = NULL)
samples69 <- samples69.trt$tr_all_wide
# print(samples69[, .N, by = .(latitude, longitude, uid_site)])

```

## note that if any variable is missing in users' dataset, the system will prompt an alert to request users to fill it for mandatory information or fill it with  for un-mandatory information under users' permission.
# 2. frequency distribution 
## count the number of trees (uid_tree) with longitude-latitude per 1 degree.

```{r, fig.width=7, fig.height=5, fig.align='center'}
dt.freq <- CFS_freq(samples69, uid_level = "uid_tree",  geo_resolution = c(1,1))
plots_freq(dt.freq, out.species = "all", caption = "CFS-TRenD1.2 samples69", out.pdf= NULL )
```


# 3. scale 
## compare the median of ring width measuement of one specific site with those of the nearby sites.


```{r, fig.width=7, fig.height=5, fig.align='center'}
# CFS_scale()
# specify the site and species to check
chk.site <- samples69[uid_site == 1 & species == "PSEUMEN",c("species", "uid_site", "site_id", "latitude", "longitude")][, .SD[1]]

dt.scale <- CFS_scale(site2chk = chk.site, ref_sites = samples69, N.nbs = 10, make.plot = list(TRUE, "CFS-TRenD1.2 samples69"))

```

# 4. quality assessment 
## Assess tree-ring measurement accuracy using a treated series based on the differences between two consecutive tree-ring measurements


```{r, fig.width=7, fig.height=5, fig.align='center'}


 samples69_long <- merge(samples69.trt$tr_all_wide[, c("uid_site", "site_id", "species", "uid_tree", "uid_sample", "uid_radius")], 
   samples69.trt$tr_all_long$tr_7_ring_widths, by = "uid_radius")
# Calculate BAI
# rename to the reserved column name
setnames(samples69_long, c("uid_sample", "year", "rw_mm"), c("SampleID", "Year" ,"RawRing"))
# check duplication
samples69_long[, .N, by = .(SampleID, Year)][N>1]
# run quality assessment procedure
qa.trt.ccf <-CFS_qa(dt.input = samples69_long)

# SampleID list
# unique(qa.trt.ccf$dt.ccf$SampleID)

# plot specific samples, figure output to screen
plots_qa(qa.trt.ccf$plot.lst, out.series = c("5"), out.pdf = NULL )

# plot all series, figures output to a pdf file
# plots_qa(qa.trt.ccf$plot.lst, out.series = "all", out.pdf = "c:/tmp/qa.pdf" )

```
# 
# 5. gamm and model selection 
## clim69 contains Frostdays for each site in samples69.
##

```{r}

samples69_clim <- merge(samples69.trt$tr_all_wide[, c("uid_site", "site_id", "species", "uid_tree", "uid_radius")], samples69.trt$tr_all_long$tr_7_ring_widths, by = "uid_radius")
# Calculate BAI
setorder(samples69_clim, uid_radius, year)
samples69_clim[, .N, by = .(uid_radius, year)][N>1]
samples69_clim[, `:=`(ageC = seq_len(.N),
  radius = cumsum(rw_mm),  # Cumulative radius (assumes RW is added each year)
          radius_prev = shift(cumsum(rw_mm), fill = 0)), by = uid_radius]  # Previous radius (shifted)

# Compute BA in cm2
samples69_clim[, ba_cm2_t_1 := pi * (radius_prev^2)/100]

samples69_clim[, bai_cm2 := pi * (radius^2 - radius_prev^2)/100]


# Drop the radius columns if not needed
samples69_clim[, c("radius", "radius_prev") := NULL]

samples69_clim <- merge(samples69_clim, clim69, by = c("site_id", "year"))

samples69_clim[, uid_tree.fac:= as.factor(as.character(uid_tree))]
setorder(samples69_clim, uid_tree, year)
# choose 1 site, ignore the first 10 years as Girardin 2022
site.i <- samples69_clim[uid_site == 10][ageC > 10]
det.i <- detrend_site(data = site.i, 
             m.candidates = c(
  "bai_cm2 ~ log(ba_cm2_t_1) + s(ageC)",
  "bai_cm2 ~ log(ba_cm2_t_1) + s(ageC) + FFD",
  "bai_cm2 ~ log(ba_cm2_t_1) + s(ageC) + s(FFD)"
  
)
)
print(summary(det.i$model$gam))
```
